```{r setup, include=FALSE}

library(data.table)
library(tidyverse)
library(readxl)
library(ggplot2)
library(stringr)
library(FSA)
library(dunn.test)
library(Epi)
library(epitools)
```


```{r, results='markup'}
dt = readRDS("data_out/dt.RDS"); dt[,.N, vaccine]; 
dt[,.N, .(record_id,vaccine)][,.N,vaccine]
```

# subsets

```{r, results='markup'}
dt.use = dt[is.na(exclude),]; dt.use[,.N, vaccine]; dt.use[,.N, .(record_id,vaccine)][,.N,vaccine]
```

Use for demographics table:
```{r, results='markup'}
dt.demo = dt.use[time_point == "Baseline"]
```

Use for treatment table:
```{r}
dt.rx = 
    melt( dt.demo
      , id.vars = c("mrn","vaccine")
      , measure.vars = c("treatment_type","treatment_type_2", "treatment_type_3")
      , variable.name = "del"
      , value.name = "treatment"
      )

dt.rx[ treatment == "",.N, vaccine]
dt.rx = dt.rx[ treatment != ""]
head(dt.rx)

dt.rx[, treatment_2 := treatment]

dt.rx[treatment %in% c("anti-CD-38 antibody","RANKL monoclonal","anti-HER antibody","anti-VEGF antibody","anti-CD-20 antibody")
      , treatment_2 := "monoclonal antibody"]

dt.rx[, treatment_2 := 
          factor(treatment_2
                 , levels = c("clinical surveillance","endocrine","kinase inhibitor","immuno- therapy","cytotoxic","monoclonal antibody","unknown")
                 , ordered = TRUE)
      ]
dt.rx[,.N,treatment_2]
```


```{r}
dt.rx.titer =
    merge(dt.rx
          , dt.use[, .(mrn,time_point,titer,titer_cat)]
          , by = "mrn")
```


# Table 2

Rows per timepoint:
```{r, results='markup'}
dt.use[time_point != "Baseline", .N , keyby = .(time_point, titer_cat)]
```

## Overall

```{r, results='markup'}
merge(
dt.use[time_point != "Baseline" & titer_cat == "positive", .(positive = .N) , keyby = .(time_point)]  ,
dt.use[time_point != "Baseline" & !is.na(titer)
       , .( denominator = .N
           , median = round( median(titer) )
           , ll     = round( quantile(titer, probs = c(0.25)) )
           , ul     = round( quantile(titer, probs = c(0.75)) )
          )
      , keyby = .(time_point) ] ,
by = c("time_point")
)[, .(time_point
      , positive = paste0( positive, "/", denominator, " (", percentage = round(100*positive/denominator), "%)" )
      , titer = paste0(median, " ( ", ll, " - ", ul, " )" ) )]
```

Compare the sero-conversation between the time-points.
```{r, results='markup'}
fisher.test(tab)
fisher.test(tab, alternative = "less")
rm(tab)
```

Compare the titer between the time-points.
```{r, results='markup'}
dat = dt.use[time_point != "Baseline" & !is.na(titer), .(time_point = factor(time_point, ordered = FALSE), titer) ]
# kruskal-Wallis
kruskal.test( formula = titer ~ time_point, data = dat )
dunn.test(x = dat$titer, g = dat$time_point)
# dunnTest(titer ~ time_point, data = dat, method="bh" ) 
```


## Vaccine
```{r, results='markup'}
merge(
dt.use[time_point != "Baseline" & titer_cat == "positive", .(positive = .N) , keyby = .(time_point, vaccine)]  ,
dt.use[time_point != "Baseline" & !is.na(titer)
       , .( denominator = .N
           , median = round( median(titer) )
           , ll     = round( quantile(titer, probs = c(0.25)) )
           , ul     = round( quantile(titer, probs = c(0.75)) )
          )
      , keyby = .(time_point, vaccine) ] ,
by = c("time_point","vaccine")
)[, .(time_point
      , vaccine
      , positive = paste0( positive, "/", denominator, " (", percentage = round(100*positive/denominator), "%)" )
      , titer = paste0(median, " ( ", ll, " - ", ul, " )" ) )]
```

Compare the sero-conversation between the vaccines stratified by time-points.
\\
1st time-point:
```{r, results='markup'}
p1 = 24; p2 = 74; d1 = 29; d2 = 92
n1 = d1 - p1; n2 = d2 - p2
tab <-
    matrix(c(p1, p2, n1, n2)
           , nrow = 2
           , dimnames = list(c("BNT162b2", "mRNA-1273"), c("positive", "negative"))
    ); tab
fisher.test(tab)
rm(tab)
```


2nd time-point:
```{r, results='markup'}
p1 = 28; p2 = 88; d1 = 30; d2 = 93
n1 = d1 - p1; n2 = d2 - p2
tab <-
    matrix(c(p1, p2, n1, n2)
           , nrow = 2
           , dimnames = list(c("BNT162b2", "mRNA-1273"), c("positive", "negative"))
    ); tab
fisher.test(tab)
rm(tab)
```

Compare the titer between the vaccines stratified by time-points.
\\
Post 1st dose
```{r, results='markup'}
dat = dt.use[time_point == "Post 1st dose" & !is.na(titer), .(vaccine = factor(vaccine, ordered = FALSE), titer) ]
kruskal.test( formula = titer ~ vaccine, data = dat )
```

Post 2nd dose
```{r, results='markup'}
dat = dt.use[time_point == "Post 2nd dose" & !is.na(titer), .(vaccine = factor(vaccine, ordered = FALSE), titer) ]
kruskal.test( formula = titer ~ vaccine, data = dat )
```



## age

```{r, results='markup'}
merge(
dt.use[time_point != "Baseline" & titer_cat == "positive", .(positive = .N) , keyby = .(time_point, age_binary)]  ,
dt.use[time_point != "Baseline" & !is.na(titer)
       , .( denominator = .N
           , median = round( median(titer) )
           , ll     = round( quantile(titer, probs = c(0.25)) )
           , ul     = round( quantile(titer, probs = c(0.75)) )
          )
      , keyby = .(time_point, age_binary) ] ,
by = c("time_point","age_binary")
)[, .(time_point
      , age_binary
      , positive = paste0( positive, "/", denominator, " (", percentage = round(100*positive/denominator), "%)" )
      , titer = paste0(median, " ( ", ll, " - ", ul, " )" ) )]
```


Compare the sero-conversation between the age stratified by time-points.
\\
1st time-point:
```{r, results='markup'}
p1 = 54; d1 = 64; p2 = 44; d2 = 57;   n1 = d1 - p1; n2 = d2 - p2
tab <- matrix(c(p1, p2, n1, n2), nrow = 2, dimnames = list(c("younger than 65", "65 and older"), c("positive", "negative"))); tab
fisher.test(tab); rm(tab)
```

2nd time-point:
```{r, results='markup'}
p1 = 64; d1 = 66; p2 = 52; d2 = 57;   n1 = d1 - p1; n2 = d2 - p2
tab <- matrix(c(p1, p2, n1, n2), nrow = 2, dimnames = list(c("younger than 65", "65 and older"), c("positive", "negative"))); tab
fisher.test(tab); rm(tab)
```

Compare the titer between the ages stratified by time-points.
\\
Post 1st dose
```{r, results='markup'}
dat = dt.use[time_point == "Post 1st dose" & !is.na(titer), .(age_binary = factor(age_binary, ordered = FALSE), titer) ]
kruskal.test( formula = titer ~ age_binary, data = dat )
```

Post 2nd dose
```{r, results='markup'}
dat = dt.use[time_point == "Post 2nd dose" & !is.na(titer), .(vaccine = factor(vaccine, ordered = FALSE), titer) ]
kruskal.test( formula = titer ~ vaccine, data = dat )
```











## sex
```{r, results='markup'}
merge(
dt.use[time_point != "Baseline" & titer_cat == "positive", .(positive = .N) , keyby = .(time_point, sex)]  ,
dt.use[time_point != "Baseline" & !is.na(titer)
       , .( denominator = .N
           , median = round( median(titer) )
           , ll     = round( quantile(titer, probs = c(0.25)) )
           , ul     = round( quantile(titer, probs = c(0.75)) )
          )
      , keyby = .(time_point, sex) ] ,
by = c("time_point","sex")
)[, .(time_point
      , sex
      , positive = paste0( positive, "/", denominator, " (", percentage = round(100*positive/denominator), "%)" )
      , titer = paste0(median, " ( ", ll, " - ", ul, " )" ) )]
```

Compare the sero-conversation between the age stratified by time-points.
\\
1st time-point:
```{r, results='markup'}
p1 = 53; d1 = 69; p2 = 45; d2 = 52;   n1 = d1 - p1; n2 = d2 - p2
tab <- matrix(c(p1, p2, n1, n2), nrow = 2, dimnames = list(c("male", "female"), c("positive", "negative"))); tab
fisher.test(tab); rm(tab)
```

2nd time-point:
```{r, results='markup'}
p1 = 64; d1 = 69; p2 = 52; d2 = 54;   n1 = d1 - p1; n2 = d2 - p2
tab <- matrix(c(p1, p2, n1, n2), nrow = 2, dimnames = list(c("male", "female"), c("positive", "negative"))); tab
fisher.test(tab); rm(tab)
```

Compare the titer between the ages stratified by time-points.
\\
Post 1st dose
```{r, results='markup'}
dat = dt.use[time_point == "Post 1st dose" & !is.na(titer), .(sex = factor(sex, ordered = FALSE), titer) ]
kruskal.test( formula = titer ~ sex, data = dat )
```

Post 2nd dose
```{r, results='markup'}
dat = dt.use[time_point == "Post 2nd dose" & !is.na(titer), .(sex = factor(sex, ordered = FALSE), titer) ]
kruskal.test( formula = titer ~ sex, data = dat )
```









## race_ethnicity
```{r, results='markup'}
merge(
dt.use[time_point != "Baseline" & titer_cat == "positive", .(positive = .N) , keyby = .(time_point, race_ethnicity)]  ,
dt.use[time_point != "Baseline" & !is.na(titer)
       , .( denominator = .N
           , median = round( median(titer) )
           , ll     = round( quantile(titer, probs = c(0.25)) )
           , ul     = round( quantile(titer, probs = c(0.75)) )
          )
      , keyby = .(time_point, race_ethnicity) ] ,
by = c("time_point","race_ethnicity")
)[, .(time_point
      , race_ethnicity
      , positive = paste0( positive, "/", denominator, " (", percentage = round(100*positive/denominator), "%)" )
      , titer = paste0(median, " ( ", ll, " - ", ul, " )" ) )]
```


Compare the sero-conversation between the race stratified by time-points.
\\
1st time-point:
```{r, results='markup'}
p1 = 79; d1 = 100; p2 = 18; d2 = 19; p3 = 1; d3 = 2;   n1 = d1 - p1; n2 = d2 - p2; n3 = d3 - p3
tab <- matrix(c(p1, p2, p3, n1, n2, n3), nrow = 3, dimnames = list(c("NHW", "Hispanic", "Black"), c("positive", "negative"))); tab
fisher.test(tab); rm(tab)
```

2nd time-point:
```{r, results='markup'}
p1 = 96; d1 = 102; p2 = 18; d2 = 18; p3 = 2; d3 = 3;   n1 = d1 - p1; n2 = d2 - p2; n3 = d3 - p3
tab <- matrix(c(p1, p2, p3, n1, n2, n3), nrow = 3, dimnames = list(c("NHW", "Hispanic", "Black"), c("positive", "negative"))); tab
fisher.test(tab); rm(tab)
```

Compare the titer between the race stratified by time-points.
\\
Post 1st dose
```{r, results='markup'}
dat = dt.use[time_point == "Post 1st dose" & !is.na(titer), .(race_ethnicity = factor(race_ethnicity, ordered = FALSE), titer) ]
kruskal.test( formula = titer ~ race_ethnicity, data = dat )
```

Post 2nd dose
```{r, results='markup'}
dat = dt.use[time_point == "Post 2nd dose" & !is.na(titer), .(race_ethnicity = factor(race_ethnicity, ordered = FALSE), titer) ]
kruskal.test( formula = titer ~ race_ethnicity, data = dat )
```







## cancer_binary
```{r, results='markup'}
merge(
dt.use[time_point != "Baseline" & titer_cat == "positive", .(positive = .N) , keyby = .(time_point, cancer_binary)]  ,
dt.use[time_point != "Baseline" & !is.na(titer)
       , .( denominator = .N
           , median = round( median(titer) )
           , ll     = round( quantile(titer, probs = c(0.25)) )
           , ul     = round( quantile(titer, probs = c(0.75)) )
          )
      , keyby = .(time_point, cancer_binary) ] ,
by = c("time_point","cancer_binary")
)[, .(time_point
      , cancer_binary
      , positive = paste0( positive, "/", denominator, " (", percentage = round(100*positive/denominator), "%)" )
      , titer = paste0(median, " ( ", ll, " - ", ul, " )" ) )]
```


Compare the sero-conversation between the cancer stratified by time-points.
\\
1st time-point:
```{r, results='markup'}
p1 = 80; d1 = 96; p2 = 18; d2 = 25;   n1 = d1 - p1; n2 = d2 - p2
tab <- matrix(c(p1, p2, n1, n2), nrow = 2, dimnames = list(c("solid tumor", "hematologic"), c("positive", "negative"))); tab
fisher.test(tab); rm(tab)
```

2nd time-point:
```{r, results='markup'}
p1 = 99; d1 = 101; p2 = 17; d2 = 22;   n1 = d1 - p1; n2 = d2 - p2
tab <- matrix(c(p1, p2, n1, n2), nrow = 2, dimnames = list(c("solid tumor", "hematologic"), c("positive", "negative"))); tab
fisher.test(tab); rm(tab)
```

Compare the titer between the cancer stratified by time-points.
\\
Post 1st dose
```{r, results='markup'}
dat = dt.use[time_point == "Post 1st dose" & !is.na(titer), .(cancer_binary = factor(cancer_binary, ordered = FALSE), titer) ]
kruskal.test( formula = titer ~ cancer_binary, data = dat )
```

Post 2nd dose
```{r, results='markup'}
dat = dt.use[time_point == "Post 2nd dose" & !is.na(titer), .(cancer_binary = factor(cancer_binary, ordered = FALSE), titer) ]
kruskal.test( formula = titer ~ cancer_binary, data = dat )
```









## treatment

```{r, results='markup'}
merge(
dt.rx.titer[time_point != "Baseline" & titer_cat == "positive", .(positive = .N) , keyby = .(time_point, treatment)]  ,
dt.rx.titer[time_point != "Baseline" & !is.na(titer)
       , .( denominator = .N
           , median = round( median(titer) )
           , ll     = round( quantile(titer, probs = c(0.25)) )
           , ul     = round( quantile(titer, probs = c(0.75)) )
          )
      , keyby = .(time_point, treatment) ] ,
by = c("time_point","treatment"), all = T
)[, .(time_point
      , treatment
      , positive = paste0( positive, "/", denominator, " (", percentage = round(100*positive/denominator), "%)" )
      , titer = paste0(median, " ( ", ll, " - ", ul, " )" ) )]
```


```{r, results='markup'}
merge(
dt.rx.titer[time_point != "Baseline" & titer_cat == "positive", .(positive = .N) , keyby = .(time_point, treatment_2)]  ,
dt.rx.titer[time_point != "Baseline" & !is.na(titer)
       , .( denominator = .N
           , median = round( median(titer) )
           , ll     = round( quantile(titer, probs = c(0.25)) )
           , ul     = round( quantile(titer, probs = c(0.75)) )
          )
      , keyby = .(time_point, treatment_2) ] ,
by = c("time_point","treatment_2"), all = T
)[, .(time_point
      , treatment_2
      , positive = paste0( positive, "/", denominator, " (", percentage = round(100*positive/denominator), "%)" )
      , titer = paste0(median, " ( ", ll, " - ", ul, " )" ) )]
```


Compare the sero-conversation between the treatment stratified by time-points.
\\
1st time-point:
```{r, results='markup'}
p1 = 3; d1 = 4; p2 = 0; d2 = 4; p3 = 1; d3 = 1; p4 = 2; d4 = 2; p5 = 4; d5 = 5; p6 = 38; d6 = 44; p7= 20; d7 = 29; p8 = 15; d8 = 16;  p9 = 11; d9 = 13;  p10 = 13; d10 = 15; 
n1 = d1 - p1; n2 = d2 - p2; n3 = d3 - p3; n4 = d4 - p4; n5 = d5 - p5; n6 = d6 - p6; n7 = d7 - p7; n8 = d8 - p8; n9 = d9 - p9; n10 = d10 - p10;
dt.rx.titer[,.N,keyby = treatment][1:10,treatment]
tab <- matrix(c(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10, n1,n2,n3,n4,n5,n6,n7,n8,n9,n10), nrow = 10
              , dimnames = list(dt.rx.titer[,.N,keyby = treatment][1:10,treatment], c("positive", "negative"))); tab
fisher.test(tab); #rm(tab)
```

```{r}
library("RVAideMemoire")
fisher.multcomp(tab, p.method = "BH")
```

2nd time-point:
```{r, results='markup'}
p1 = 3; d1 = 3; p2 = 0; d2 = 4; p3 = 1; d3 = 1; p4 = 1; d4 = 1; p5 = 5; d5 = 5; p6 = 44; d6 = 45; p7= 28; d7 = 30; p8 = 18; d8 = 18;  p9 = 13; d9 = 14;  p10 = 12; d10 = 13; 
n1 = d1 - p1; n2 = d2 - p2; n3 = d3 - p3; n4 = d4 - p4; n5 = d5 - p5; n6 = d6 - p6; n7 = d7 - p7; n8 = d8 - p8; n9 = d9 - p9; n10 = d10 - p10;
dt.rx.titer[,.N,keyby = treatment][1:10,treatment]
tab <- matrix(c(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10, n1,n2,n3,n4,n5,n6,n7,n8,n9,n10), nrow = 10
              , dimnames = list(dt.rx.titer[,.N,keyby = treatment][1:10,treatment], c("positive", "negative"))); tab
fisher.test(tab); #rm(tab)
```

```{r}
library("RVAideMemoire")
fisher.multcomp(tab, p.method = "BH")
```

Compare the titer between the treatment stratified by time-points.
\\
Post 1st dose
```{r, results='markup'}
dat = dt.rx.titer[time_point == "Post 1st dose" & !is.na(titer), .(treatment= factor(treatment, ordered = FALSE), titer) ]
kruskal.test( formula = titer ~ treatment, data = dat )
dunnTest(titer ~ treatment, data = dat, method="bh" ) 
data.table(dunnTest(titer ~ treatment, data = dat, method="bh" )$res)[P.adj < 0.05, .(Comparison, P.adj)]
```

Post 2nd dose
```{r, results='markup'}
dat = dt.rx.titer[time_point == "Post 2nd dose" & !is.na(titer), .(treatment= factor(treatment, ordered = FALSE), titer) ]
kruskal.test( formula = titer ~ treatment, data = dat )
dunnTest(titer ~ treatment, data = dat, method="bh" ) 
data.table(dunnTest(titer ~ treatment, data = dat, method="bh" )$res)[P.adj < 0.05, .(Comparison, P.adj)]
```

## Treatment 2

Compare the sero-conversation between the treatment stratified by time-points.
\\
1st time-point:
```{r, results='markup'}
p1 = 10; d1 = 16; p6 = 38; d6 = 44; p7= 20; d7 = 29; p8 = 15; d8 = 16;  p9 = 11; d9 = 13;  p10 = 13; d10 = 15; 
n1 = d1 - p1; n6 = d6 - p6; n7 = d7 - p7; n8 = d8 - p8; n9 = d9 - p9; n10 = d10 - p10;
dt.rx.titer[,.N,keyby = treatment][c(1,6:10),treatment]
tab <- matrix(c(p1,p6,p7,p8,p9,p10, n1,n6,n7,n8,n9,n10), nrow = 6
              , dimnames = list(dt.rx.titer[,.N,keyby = treatment][c(1,6:10),treatment], c("positive", "negative"))); tab
fisher.test(tab); #rm(tab)
```

```{r}
library("RVAideMemoire")
fisher.multcomp(tab, p.method = "BH")
```

2nd time-point:
```{r, results='markup'}
p1 = 10; d1 = 14; p6 = 44; d6 = 45; p7= 28; d7 = 30; p8 = 18; d8 = 18;  p9 = 13; d9 = 14;  p10 = 12; d10 = 13; 
n1 = d1 - p1; n6 = d6 - p6; n7 = d7 - p7; n8 = d8 - p8; n9 = d9 - p9; n10 = d10 - p10;
dt.rx.titer[,.N,keyby = treatment][c(1,6:10),treatment]
tab <- matrix(c(p1,p6,p7,p8,p9,p10, n1,n6,n7,n8,n9,n10), nrow = 6
              , dimnames = list(dt.rx.titer[,.N,keyby = treatment][c(1,6:10),treatment], c("positive", "negative"))); tab
fisher.test(tab); #rm(tab)
```

```{r}
library("RVAideMemoire")
fisher.multcomp(tab, p.method = "BH")
```

Compare the titer between the treatment stratified by time-points.
\\
Post 1st dose
```{r, results='markup'}
dat = dt.rx.titer[time_point == "Post 1st dose" & !is.na(titer), .(treatment_2= factor(treatment_2, ordered = FALSE), titer) ]
kruskal.test( formula = titer ~ treatment_2, data = dat )
dunnTest(titer ~ treatment_2, data = dat, method="bh" ) 
data.table(dunnTest(titer ~ treatment_2, data = dat, method="bh" )$res)[P.adj < 0.05, .(Comparison, P.adj)]
```

Post 2nd dose
```{r, results='markup'}
dat = dt.rx.titer[time_point == "Post 2nd dose" & !is.na(titer), .(treatment_2= factor(treatment_2, ordered = FALSE), titer) ]
kruskal.test( formula = titer ~ treatment_2, data = dat )
dunnTest(titer ~ treatment_2, data = dat, method="bh" ) 
data.table(dunnTest(titer ~ treatment_2, data = dat, method="bh" )$res)[P.adj < 0.05, .(Comparison, P.adj)]
```










# Table 1


## Age

```{r, results='markup'}
stat.table(
	index = list( "vaccine" = vaccine ),
	contents=list( "freq" = count(), "mean" = mean(age), "std.dev" = sd(age), "median" = median(age) ),
	data=dt.demo,
    margins=c(TRUE)
	)
```

__Interquartile range__

Overall
```{r, results='markup'}
quantile(dt.demo[,age], probs = c(0.25,0.5,0.75))
```

mRNA-1273
```{r, results='markup'}
quantile(dt.demo[vaccine == "mRNA-1273",age], probs = c(0.25,0.5,0.75))
```

BNT162b2
```{r, results='markup'}
quantile(dt.demo[vaccine == "BNT162b2",age], probs = c(0.25,0.5,0.75))
```

## Gender

```{r, results='markup'}
stat.table(
	index = list( "gender" = sex, "vaccine" = vaccine ),
	contents=list( count(), percent(sex) ),
	data=dt.demo,
    margins=c(TRUE, TRUE)
	)
```

## Race/Ethnicity

```{r, results='markup'}
stat.table(
	index = list( "Race/Ethnicity" = race_ethnicity, "vaccine" = vaccine ),
	contents=list( count(), percent(race_ethnicity) ),
	data=dt.demo,
    margins=c(TRUE, TRUE)
	)
```


## Response measurement time range

```{r, results='markup'}
stat.table(
	index = list( "vaccine" = vaccine ),
	contents=list( "freq" = count(), "mean" = mean(days_to_titer), "std.dev" = sd(days_to_titer), "median" = median(days_to_titer) ),
	data=dt.use[time_point == "Post 2nd dose",],
    margins=c(TRUE)
	)
```

__Interquartile range__

Overall
```{r, results='markup'}
quantile(dt.use[time_point == "Post 2nd dose", days_to_titer], probs = c(0.25,0.5,0.75))
```

mRNA-1273
```{r, results='markup'}
quantile(dt.use[time_point == "Post 2nd dose" & vaccine == "mRNA-1273", days_to_titer], probs = c(0.25,0.5,0.75))
```

BNT162b2
```{r, results='markup'}
quantile(dt.use[time_point == "Post 2nd dose" & vaccine == "BNT162b2",days_to_titer], probs = c(0.25,0.5,0.75))
```


## Malignancy Grouper

```{r, results='markup'}
stat.table(
	index = list( "malignancy" = cancer_binary, "vaccine" = vaccine ),
	contents=list( count(), percent(cancer_binary) ),
	data=dt.demo,
    margins=c(TRUE, TRUE)
	)
```

## Malignancy System

```{r, results='markup'}
stat.table(
	index = list( "malignancy system" = cancer_system, "vaccine" = vaccine ),
	contents=list( count(), percent(cancer_system) ),
	data=dt.demo,
    margins=c(TRUE, TRUE)
	)
```

## Treatment




```{r, results='markup'}
stat.table(
	index = list( "treatment" = treatment, "vaccine" = vaccine ),
	contents=list( count(), percent(treatment) ),
	data=dt.rx,
    margins=c(TRUE, TRUE)
	)
```


```{r, results='markup'}
stat.table(
	index = list( "treatment" = treatment_2, "vaccine" = vaccine ),
	contents=list( count(), percent(treatment) ),
	data=dt.rx,
    margins=c(TRUE, TRUE)
	)
```


# Graphs

```{r}
c.2colors = c("#aac872", "#f5af5c")
c.2greys  = c("#aac872", "#de425b")

c.2colors = c("grey75","grey75")
c.2greys = c("grey70","grey30")
```

## Vaccine response overall

```{r}
dat = dt.use[time_point != "Baseline" & !is.na(titer)]
dat[, time_point := factor(time_point, levels = c("Post 1st dose","Post 2nd dose"), labels = c("Timepoint 1","Timepoint 2"))]
str(dat$time_point)
```

```{r}
gg.1 =
    ggplot(dat, aes(x=time_point, y=titer, fill=time_point, color = time_point)) +
    # geom_violin(width = 1) +
    geom_jitter(width = 0.25, height = 0.25, size = 1, alpha = 0.3, color = "grey50", fill = "grey 80") +
    geom_boxplot(width=0.25, color="black", alpha=0, outlier.colour = NA) +
    #scale_fill_viridis(discrete = TRUE) +
    theme_classic() +
    theme(
      legend.position="none",
      text = element_text(size=10)
      # plot.margin=unit(c(1,0.2,0.2,0.2),"cm")
      # axis.title.x = element_text(margin = unit(c(0.5,0,0,0),"cm"))
    ) +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 11)) +
    scale_y_continuous(breaks=seq(0,2500,500)) +
    scale_color_manual(values=c.2colors) +
    scale_fill_manual(values=c.2colors) +
    # ggtitle("Response") +
    xlab(label = NULL) +
    ylab("Anti-SARS-CoV-2 S (U/mL)")

gg.1 + 
    annotate("segment", x = 1, xend = 2, y = 2600, yend = 2600, colour = "grey 30", size = 0.5 ) +
    annotate("text", x = 1.5, y = 2700, label = "paste(italic(p), \"< 0.001\")", parse = TRUE, size = 2) +
    annotate("text", x = 1, y = - 100, label = "N = 121", size = 2) +
    annotate("text", x = 2, y = - 100, label = "N = 123", size = 2)
ggsave( "01_overall.pdf", path = "./plot", width = 3, height = 3, units = "in", dpi = 500)
rm(gg.1)
```



## Vaccine response by vaccine
```{r}
gg.1 =
    ggplot(dat, aes(x=vaccine, y=titer, fill=vaccine, color = vaccine)) +
    # geom_violin(width = 0.9) +
    geom_jitter(width = 0.25, height = 0.25, size = 1, alpha = 0.3, color = "grey50", fill = "grey 80") +
    geom_boxplot(width=0.25, color="black", alpha=0, outlier.colour = NA) +
    #scale_fill_viridis(discrete = TRUE) +
    theme_classic() +
    theme(
      legend.position="none",
      text = element_text(size=9)
    ) +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 4)) +
    scale_y_continuous(breaks=seq(0,2500,500)) +
    scale_color_manual(values=c.2colors) +
    scale_fill_manual(values=c.2colors) +
    # ggtitle("Response by vaccine type") +
    xlab(label = NULL) +
    ylab("Anti-SARS-CoV-2 S (U/mL)") + 
    facet_wrap( ~ time_point # , "    p=", pvalue
                #, scales = "free"
                #, strip.position = "left"
                , ncol = 2)
gg.1 +
    annotate("segment", x = 1, xend = 2, y = 2600, yend = 2600, colour = "grey 30", size = 0.5 ) +
    annotate("text", x = 1.5, y = 2700, label = "paste(italic(p), \"=      \")", parse = TRUE, size = 2) +
    annotate("text", x = 1, y = - 100, label = "N =      ", size = 2) +
    annotate("text", x = 2, y = - 100, label = "N =      ", size = 2)
ggsave( "02_vaccine_type.pdf", path = "./plot", width = 3, height = 3, units = "in", dpi = 500)
rm(gg.1)
```


## Vaccine response by gender
```{r}
gg.1 =
    ggplot(dat, aes(x=sex, y=titer, fill=sex, color = sex)) +
    # geom_violin(width = 0.9) +
    geom_jitter(width = 0.25, height = 0.25, size = 1, alpha = 0.3, color = "grey50", fill = "grey 80") +
    geom_boxplot(width=0.25, color="black", alpha=0, outlier.colour = NA) +
    #scale_fill_viridis(discrete = TRUE) +
    theme_classic() +
    theme(
      legend.position="none",
      text = element_text(size=10)
    ) +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 11)) +
    scale_y_continuous(breaks=seq(0,2500,500)) +
    scale_color_manual(values=c.2colors) +
    scale_fill_manual(values=c.2colors) +
    # ggtitle("Response by gender") +
    xlab(label = NULL) +
    ylab("Anti-SARS-CoV-2 S (U/mL)") + 
    facet_wrap( ~ time_point # , "    p=", pvalue
                #, scales = "free"
                #, strip.position = "left"
                , ncol = 2)
gg.1 +
    annotate("segment", x = 1, xend = 2, y = 2600, yend = 2600, colour = "grey 30", size = 0.5 ) +
    annotate("text", x = 1.5, y = 2700, label = "paste(italic(p), \"=      \")", parse = TRUE, size = 2) +
    annotate("text", x = 1, y = - 100, label = "N =      ", size = 2) +
    annotate("text", x = 2, y = - 100, label = "N =      ", size = 2)
ggsave( "03_gender.pdf", path = "./plot", width = 3, height = 3, units = "in", dpi = 500)
rm(gg.1)
```


## Vaccine response by cancer
```{r}
gg.1 =
    ggplot(dat, aes(x=cancer_binary, y=titer, fill=cancer_binary, color=cancer_binary)) +
    # geom_violin(width = 0.9) +
    geom_jitter(width = 0.25, height = 0.25, size = 1, alpha = 0.3, color = "grey50", fill = "grey 80") +
    geom_boxplot(width=0.25, color="black", alpha=0, outlier.colour = NA) +
    #scale_fill_viridis(discrete = TRUE) +
    theme_classic() +
    theme(
      legend.position="none",
      text = element_text(size=10)
    ) +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 9)) +
    scale_y_continuous(breaks=seq(0,2500,500)) +
    scale_color_manual(values=c.2colors) +
    scale_fill_manual(values=c.2colors) +
    # ggtitle("Response by malignancy") +
    xlab(label = NULL) +
    ylab("Anti-SARS-CoV-2 S (U/mL)") + 
    facet_wrap( ~ time_point # , "    p=", pvalue
                #, scales = "free"
                #, strip.position = "left"
                , ncol = 2)
gg.1 +
    annotate("segment", x = 1, xend = 2, y = 2600, yend = 2600, colour = "grey 30", size = 0.5 ) +
    annotate("text", x = 1.5, y = 2700, label = "paste(italic(p), \"=      \")", parse = TRUE, size = 2) +
    annotate("text", x = 1, y = - 100, label = "N =      ", size = 2) +
    annotate("text", x = 2, y = - 100, label = "N =      ", size = 2)
ggsave( "04_tumor.pdf", path = "./plot", width = 3, height = 3, units = "in", dpi = 500)
rm(gg.1)
```


## Vaccine response by treatment


```{r}
dat = dt.rx.titer[time_point == "Post 2nd dose" & treatment_2 != "unknown" & !is.na(titer)]
gg.1 =
    ggplot(dat, aes(x=treatment_2, y=titer, fill=treatment_2, color = treatment_2)) +
    # geom_violin(width = 0.9) +
    geom_jitter(width = 0.25, height = 0.25, size = 1, alpha = 0.3, color = "grey50", fill = "grey 80") +
    geom_boxplot(width=0.25, color="black", alpha=0, outlier.colour = NA) +
    #scale_fill_viridis(discrete = TRUE) +
    theme_classic() +
    theme(
      legend.position="none",
      text = element_text(size=10)
    ) +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 11)) +
    scale_y_continuous(breaks=seq(0,2500,500)) +
    # scale_color_manual(values=c("#00876c","#61a96e","#aac872","#f7e382","#f5af5c","#ea784d")) +
    # scale_fill_manual(values=c("#00876c","#61a96e","#aac872","#f7e382","#f5af5c","#ea784d")) +
    # ggtitle("Response by treatment") +
    xlab("Anti-Cancer Therapy") +
    ylab("Anti-SARS-CoV-2 S (U/mL)") 

gg.1 +
    annotate("segment", x = 1, xend = 6, y = 3500, yend = 3500, colour = "grey 30", size = 0.5 ) +
    annotate("text", x = 5.7, y = 3600, label = "paste(italic(p), \"=      \")", parse = TRUE, size = 2) +
    annotate("segment", x = 2, xend = 6, y = 3300, yend = 3300, colour = "grey 30", size = 0.5 ) +
    annotate("text", x = 5.7, y = 3400, label = "paste(italic(p), \"=      \")", parse = TRUE, size = 2) +
    
    annotate("segment", x = 1, xend = 5, y = 3100, yend = 3100, colour = "grey 30", size = 0.5 ) +
    annotate("text", x = 4.7, y = 3200, label = "paste(italic(p), \"=      \")", parse = TRUE, size = 2) +
    annotate("segment", x = 2, xend = 5, y = 2900, yend = 2900, colour = "grey 30", size = 0.5 ) +
    annotate("text", x = 4.7, y = 3000, label = "paste(italic(p), \"=      \")", parse = TRUE, size = 2) +
    
    annotate("text", x = 1, y = - 100, label = "N =      ", size = 2) +
    annotate("text", x = 2, y = - 100, label = "N =      ", size = 2) +
    annotate("text", x = 3, y = - 100, label = "N =      ", size = 2) +
    annotate("text", x = 4, y = - 100, label = "N =      ", size = 2) +
    annotate("text", x = 5, y = - 100, label = "N =      ", size = 2) +
    annotate("text", x = 6, y = - 100, label = "N =      ", size = 2)
ggsave( "05_treatment.pdf", path = "./plot", width = 5, height = 4, units = "in", dpi = 300)
rm(gg.1)
```





## individual vaccine response

```{r}
dat = dt.use[!is.na(titer)]
dat[, time_point := factor(time_point, levels = c("Baseline","Post 1st dose","Post 2nd dose"), labels = c("Baseline","Timepoint 1","Timepoint 2"))]
# dat[,.N,.(mrn)][N==3,.N]
# c.mrn.alltp = dat[,.N,.(mrn)][N==3,mrn]
# dat = dat[mrn %in% c.mrn.alltp,]; dat[,.N]
str(dat$time_point)
```

```{r}
gg.1 =
    ggplot(dat, aes(x = time_point, y = titer, color = vaccine, shape = vaccine)) +
    geom_point(aes(fill = vaccine), size = 1, alpha = 0.3) +
    geom_line(aes(group = record_id, linetype = vaccine), alpha = 0.5) +
    theme_classic() +
    theme( legend.position = c(0.25, 0.85)
           , text = element_text(size=9)
           , legend.text = element_text(size=7)
           #, legend.text = element_text(size = 7, margin = margin( t=0, r=0, b=0, l=0, unit="in" ) )
           , legend.spacing.y = unit(0,'in')
           ) +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 11)) +
    scale_color_manual(values=c.2greys) +
    scale_fill_manual(values=c.2greys) +
    # ggtitle("Response for each participant") +
    xlab(label = NULL) +
    ylab("Anti-SARS-CoV-2 S (U/mL)")

gg.1
ggsave( "06_participant.pdf", path = "./plot", width = 3, height = 3, units = "in", dpi = 500)
rm(gg.1)
```
